---
title: "Group 4 Script"
author: "Group 4 - Cindy Fajardo, Jesse G Jones, Avetis Mishegyan, Helia Manesh, Caroline Solis"
date: "2023-04-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sources & Credits! - Click the Links Below

##### [Tidy Tuesday Bird FeederWatch - 2023-01-10](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-10/readme.md)

##### [Data Source - FeederWatch](https://feederwatch.org/explore/raw-dataset-requests/)

##### [Article - Over 30 Years of Standardized Bird Counts at Supplementary Feeding Stations in North America: A Citizen Science Data Report for Project FeederWatch](https://www.frontiersin.org/articles/10.3389/fevo.2021.619682/full)

##### [TidyTuesday Repository](https://github.com/rfordatascience/tidytuesday)

##### [Webpage with Canadian Provinces geojson - opendatasoft](https://data.opendatasoft.com/explore/dataset/georef-canada-province%40public/export/?disjunctive.prov_name_en)

##### [Canadian Provinces gejson Data Source - Statistics Canada](https://www.statcan.gc.ca/en/reference/licence) - Source: Statistics Canada, referenced 2023-04-12. Reproduced and distributed on an "as is" basis with the permission of Statistics Canada.

##### [Webpage with US States geojson - rstudio.github.io](https://rstudio.github.io/leaflet/choropleths.html)

##### [US States geojson](https://rstudio.github.io/leaflet/json/us-states.geojson)

## Links to Data Dictionaries

##### [Data Dictionary 1](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-10/readme.md#pfw_2021_publiccsv)

##### [Data Dictionary 2](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-10/readme.md#pfw_count_site_data_public_2021csv)

```{r Libraries, echo=FALSE}
library(tidyverse)
library(here)
library(leaflet)
library(geojson) #for loading in geojson files; allows us to view states and province names & borders
library(rgeos)
library(readr)
```

## Reading in Our Data

```{r Reading in Our Data}
ca_provinces <- geojsonio::geojson_read(here("data", "provinces.geojson"), what="sp") #read in the geographical and spatial data for Canadian Provinces 

us_states <- geojsonio::geojson_read("https://rstudio.github.io/leaflet/json/us-states.geojson", what = "sp") #read in the geographical and spatial data for U.S. states 

feederwatch <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-10/PFW_2021_public.csv') #read in the feederwatch data from the TidyTuesday Github repository
site_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-10/PFW_count_site_data_public_2021.csv') #read in the site data from the TidyTuesday Github repository
SpeciesCode <- read_csv(here("data","SpeciesCode.csv")) #read in Cindy's SpeciesCode csv file
```

# Avetis' Section
```{r}
### Data Analysis ######
map <- leaflet() %>%
  addTiles() %>% # function adds default OpenStreetMap tile layer (ie. basemap)
  addProviderTiles(providers$OpenStreetMap, group = "OpenStreetMap") %>% # function adds third-party tiles (ie. basemaps)
  addProviderTiles(providers$CartoDB.DarkMatter, group = "CartoDB.DarkMatter") %>% # providers$ argument names third-party basemap
  addProviderTiles(providers$NASAGIBS.ViirsEarthAtNight2012, group = "NASAGIBS.ViirsEarthAtNight2012") %>% # group argument is the name of the group the newly created layers belong to
  addProviderTiles(providers$Esri.WorldImagery, group = "Esri.WorldImagery") %>% 
  addLayersControl(baseGroups = c("OpenStreetMap", # function adds layer controls to hide/show map layers,
                                  # and baseGroups argument creates character vector where each element is the name of a group
                                  "CartoDB.DarkMatter",
                                  "NASAGIBS.ViirsEarthAtNight2012",
                                  "Esri.WorldImagery"),
                   position = "topleft") %>%
  addControl(html = "#TidyTuesday(2023-01-10) - Bird FeederWatch Data", # function adds text box, and html argument displays text
             position = "bottomright") # position argument moves text box to any corner
map # prints map
```

# Cindy's Section

## Sites used
[Leaflet Source](https://rstudio.github.io/leaflet/map_widget.html)
This is a very good source to use to help you work your way through the leaflet package

## Basic leaflet commands to learn
In this section we will create a very simple map to help visualize what leaflet does. The functions that are used are leaflet(), addTiles(), and addMarkers().
```{r, Practice}
m <- leaflet() %>% #step 1. Create a map widget by calling leaflet()
  addTiles() %>%  # Add default OpenStreetMap map tiles, Step 2. Add layers (i.e., features) to the map by using layer functions (e.g. addTiles, addMarkers, addPolygons) to modify the map widget
  addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R") # Repeat step 2 as desired. When using markers you must must use longitude and latitude in order for the markers to be placed on the map.
m  # Print the map
```


## The Formula Interface
The arguments of all layer functions can take normal R objects, such as a numeric vector for the lat argument, or a character vector of colors for the color argument. They can also take a one-sided formula, in which case the formula will be evaluated using the data argument as the environment. For example, ~ x means the variable x in the data object, and you can write arbitrary expressions on the right-hand side, e.g., ~ sqrt(x + 1)
```{r}
m = leaflet() %>% addTiles()
df = data.frame(
  lat = rnorm(100),
  lng = rnorm(100),
  size = runif(100, 5, 20),
  color = sample(colors(), 100)
)
m = leaflet(df) %>% addTiles()
m %>% addCircleMarkers(radius = ~size, color = ~color, fill = FALSE)
m %>% addCircleMarkers(radius = runif(100, 4, 10), color = c('red'))
```


## Data Wrangling
In order to create our final map, we first needed to wrangle our data. We did a full join of feeder watch data and site data. We then did a full join of the new fully joined data named **birds_joined** with the species code data and selected columns loc_id, latitude, longtitude, subnational1_code, Month, Day, Year, species_cide, how_many, sci_name, order, family.
```{r}
feed_site <- full_join(feederwatch, site_data)
birds_joined <- full_join(feed_site, SpeciesCode) %>% 
  select(loc_id, latitude, longitude, subnational1_code, Month, Day, Year, species_code, how_many, sci_name, order, family)
birds_joined$date <- paste(birds_joined$Month, birds_joined$Day, birds_joined$Year, sep = "/") # the paste function is an R based command that will merge columns together and you can separate the values with a character, like "\", sourced from (https://datasharkie.com/how-to-merge-two-columns-in-r/) 
```

## Working with Cluster Markers and adding buttons.
When adding a large amounts of markers on your map, you can cluster them together using Leaflet.markercluster plug-in. To eneable the plug-in, you can use options like **clusterOptions**. The map becomes interactive, zooming in as you select a cluster, creating smaller clusters until you reach the individual markers.

When working with clusters, you can also add buttons to your map to freeze and unfreeze your clusters. 
```{r}
leaflet(birds_joined) %>% addTiles() %>% addMarkers(
  clusterOptions = markerClusterOptions(), #here you being to use the cluster option plug in
    clusterId = "birdsCluster",
  label = ~sci_name) %>% # here you use  formula interface by making the label the scientific name of the bird that was seen.
  addEasyButton(easyButton(  
    states = list(
      easyButtonState( 
        stateName="unfrozen-markers",
        icon="ion-toggle", 
        title="Freeze Clusters",
        onClick = JS(" 
          function(btn, map) {
            var clusterManager =
              map.layerManager.getLayer('cluster', 'birdsCluster');
            clusterManager.freezeAtZoom();
            btn.state('frozen-markers');
          }")
      ),
      easyButtonState( 
        stateName="frozen-markers",
        icon="ion-toggle-filled", 
        title="UnFreeze Clusters",
        onClick = JS(" 
          function(btn, map) {
            var clusterManager =
              map.layerManager.getLayer('cluster', 'birdsCluster');
            clusterManager.unfreeze();
            btn.state('unfrozen-markers');
          }")
      )
    )
  ))
```


# Jesse's Section

## Assigning objects for use of .geojson spatial data

```{r Data Wrangling, echo=TRUE, warning=FLASE, error=FALSE}


state_labels <- us_states$name #create an object that has uses the name column from your us_states geojson file; this will be used for our labels

province_labels <- ca_provinces$prov_name_en #this object is for our Canadian provinces labels; uses the "prov_name_en" column from the ca_provinces label
```

## Use addPolygon to outline geographical areas

```{r}

state_labels <- us_states$name

leaflet(birds_joined) %>% addTiles() %>% addMarkers(
  clusterOptions = markerClusterOptions(),
  clusterId = "birdsCluster",
  label = ~sci_name) %>%
  addEasyButton(easyButton(
    states = list(
      easyButtonState(
        stateName="unfrozen-markers",
        icon="ion-toggle",
        title="Freeze Clusters",
        onClick = JS("
          function(btn, map) {
            var clusterManager =
              map.layerManager.getLayer('cluster', 'birdsCluster');
            clusterManager.freezeAtZoom();
            btn.state('frozen-markers');
          }")
      ),
      easyButtonState(
        stateName="frozen-markers",
        icon="ion-toggle-filled",
        title="UnFreeze Clusters",
        onClick = JS("
          function(btn, map) {
            var clusterManager =
              map.layerManager.getLayer('cluster', 'birdsCluster');
            clusterManager.unfreeze();
            btn.state('unfrozen-markers');
          }")
      )
    )
  )) %>% addPolygons(data=us_states, #use data from the us_states object (geojson)
              color="lavenderblush", #add an outline color for each state border
              weight=2, #set border weight to 2
              opacity=1, #set opacity to 1 (full visibility)
              dashArray = 2.5, #give your state borders dashes
              smoothFactor = 1, #smooth out the edges!
              fillColor = "mediumslateblue", #add a fill colr to the states
              fillOpacity = 0.6) #make the fill opacity .60 (visibility is 60%)
```

## Let's add highlights & labels!

```{r}
leaflet(birds_joined) %>% addTiles() %>% addMarkers( 
  clusterOptions = markerClusterOptions(),
  clusterId = "birdsCluster",
  label = ~sci_name) %>%
  addEasyButton(easyButton(
    states = list(
      easyButtonState(
        stateName="unfrozen-markers",
        icon="ion-toggle",
        title="Freeze Clusters",
        onClick = JS("
          function(btn, map) {
            var clusterManager =
              map.layerManager.getLayer('cluster', 'birdsCluster');
            clusterManager.freezeAtZoom();
            btn.state('frozen-markers');
          }")
      ),
      easyButtonState(
        stateName="frozen-markers",
        icon="ion-toggle-filled",
        title="UnFreeze Clusters",
        onClick = JS("
          function(btn, map) {
            var clusterManager =
              map.layerManager.getLayer('cluster', 'birdsCluster');
            clusterManager.unfreeze();
            btn.state('unfrozen-markers');
          }")
      )
    )
  )) %>% 
  addPolygons(data=us_states, #use data from the us_states object (geojson)
              color="lavenderblush", #add an outline color for each state border
              weight=2, #set border weight to 2
              opacity=1, #set opacity to 1 (full visibility)
              dashArray = 2.5, #give your state borders dashes
              smoothFactor = 1, #smooth out the edges!
              fillColor = "mediumslateblue", #add a fill colr to the states
              fillOpacity = 0.6, #make the fill opacity .60 (visibility is 60%)
              highlightOptions = highlightOptions( #add highlights while hovering over each state
                weight=3.5, #set the highlight border weight to 3.5
                color="lavender", #set the highlight border color to lavender
                fillOpacity = .5, #make the fill opacity 0.5 when hovering
                dashArray = 0, #have no highlight dashes when hovering
                bringToFront = TRUE), #bring your highlights to the front layer
              label=state_labels) #label your states the name of the state
```

## Adding Canadian Provinces & ProviderTiles...

```{r Leaflet Map}

province_labels <- ca_provinces$prov_name_en

leaflet(birds_joined) %>% # function adds default OpenStreetMap tile layer (ie. basemap)
  addProviderTiles(providers$OpenStreetMap, group = "OpenStreetMap") %>% # function adds third-party tiles (ie. basemaps)
  addProviderTiles(providers$CartoDB.DarkMatter, group = "CartoDB.DarkMatter") %>% # providers$ argument names third-party basemap
  addProviderTiles(providers$NASAGIBS.ViirsEarthAtNight2012, group = "NASAGIBS.ViirsEarthAtNight2012") %>% # group argument is the name of the group the newly created layers belong to
  addProviderTiles(providers$Esri.WorldImagery, group = "Esri.WorldImagery") %>% 
  addLayersControl(baseGroups = c("OpenStreetMap", # function adds layer controls to hide/show map layers,
                                  # and baseGroups argument creates character vector where each element is the name of a group
                                  "CartoDB.DarkMatter",
                                  "NASAGIBS.ViirsEarthAtNight2012",
                                  "Esri.WorldImagery"),
                   position = "topleft") %>%
  addControl(html = "#TidyTuesday(2023-01-10) - Bird FeederWatch Data", # function adds text box, and html argument displays text
             position = "bottomright")%>%  # position argument moves text box to any corner 
  addMarkers(
  clusterOptions = markerClusterOptions(),
  clusterId = "birdsCluster",
  label = ~sci_name) %>%
  addEasyButton(easyButton(
    states = list(
      easyButtonState(
        stateName="unfrozen-markers",
        icon="ion-toggle",
        title="Freeze Clusters",
        onClick = JS("
          function(btn, map) {
            var clusterManager =
              map.layerManager.getLayer('cluster', 'birdsCluster');
            clusterManager.freezeAtZoom();
            btn.state('frozen-markers');
          }")
      ),
      easyButtonState(
        stateName="frozen-markers",
        icon="ion-toggle-filled",
        title="UnFreeze Clusters",
        onClick = JS("
          function(btn, map) {
            var clusterManager =
              map.layerManager.getLayer('cluster', 'birdsCluster');
            clusterManager.unfreeze();
            btn.state('unfrozen-markers');
          }")
      )
    )
  )) %>% 
  addPolygons(data=us_states,
              color="lavenderblush",
              weight=2,
              opacity=1,
              dashArray = 2.5,
              smoothFactor = 1,
              fillColor = "mediumslateblue",
              fillOpacity = 0.6,
              highlightOptions = highlightOptions(
                weight=3.5,
                color="lavender",
                fillOpacity = .5,
                dashArray = 0,
                bringToFront = TRUE),
              label=state_labels) %>% 
  addPolygons(data=ca_provinces, #use the data from Canadian provinces geojson spatial data
              color="lavenderblush",
              weight=2,
              opacity=1,
              dashArray = 2.5,
              smoothFactor = 1,
              fillColor = "coral", #let's use a different color for Canada...
              fillOpacity = 0.6,
              highlightOptions = highlightOptions(
                weight=3.5,
                color="salmon", #and have a different highlight border color
                fillOpacity = .5,
                dashArray = 0,
                bringToFront = TRUE),
              label=province_labels) #use the Canadian province labels

```
